The Impact of Extreme Weather Events on Property Insurance Pricing

### 2023-12-09 manuscript/manuscript.tex
double spaced
refs corrected
fixed using peer review, added more figures

@@

\section{Introduction}
\label{sec:intro}
In recent years, weather catastrophes have become more frequent and severe, which poses a potential threat to homeowners' safety as 
well as financial impacts. As the frequency and severity of disasters increase, the significance of the ability of homeowners to 
protect their homes using property insurance also increases. This research will focus on exploring the relationship between the 
frequency and severity of weather catastrophes and the pricing of property insurance.

Climate disasters have become more frequent and severe in recent years. These incidents not only put homeowners' safety at risk but 
also have significant economic consequences. As weather-related risks increase in frequency and severity, the ability for homeowners to 
protect their valuable assets through property insurance has also become increasingly important This paper aims to explore the 
relationship between the frequency and severity of weather hazards and property insurance policies and prices.

The importance of this study is the fact that weather disasters can carry profound amount of consequences, which affects the individual 
homeowners and the insurance industry as a whole. Homeowners are the ones who typically bear the financial burden of disasters, and face 
the finance of all the insurance premiums, deductibles, and coverage limitations at the same time. On the other hand, the insurance 
company is tasked to manage the increased risks associated with these events, which can also have a significant impact on their financial 
stability.

While previous research has examined different aspects of this complex issue, including the impact of weather risks on insurance 
prices, there are much more waiting to be discovered. Our research is based on valuable insights into the value of previous work 
\cite{hurricaneco} aimed at assets that have provided insurance pricing of weather frequency and severity. It provides a general 
understanding of how frequency and severity affect the pricing of property insurance. We are also trying to identify effective 
strategies that can reduce the financial burden of homeowners in the face of these major disasters.

This research has used several methods, including statistical analysis and data science methods. To achieve our goals, we will draw 
data from reliable sources, such as the Insurance Information Agency, National Insurance Centers for Environmental Informations, and 
Aon's 2023 Weather, Climate and Disaster Investigation Report. This study is not only timely but critically important at a time of 
increasing weather-related risks, with the goal of contributing to homeowner well-being and the stability of the property insurance 
market.

@@

@@

@@

@@

\section{Results}
\label{sec:resu}
\subsection{Summary Statistics}
Table 4 \ref{tab:summary} presents summary statistics for the premium and disaster data. The disaster measures show substantial variability, 
highlighting the irregular nature of extremes. Hurricane and flooding perils account for the largest share of overall cost and 
insured losses.

\begin{table}[h]
    \label{tab:summary}
    \centering
    \begin{tabular}{|l|c|c|c|}
        \hline
        & Mean & SD & Range \\
        \hline
        Premium & \$\num{959.2} & \$\num{238.5208} & \$\num{536} - \$\num{1311} \\
        Number of Disasters & MeanDisaster & SDDisaster & MinDisaster - MaxDisaster \\
        Total Cost (\$) & MeanCost & SDCost & MinCost - MaxCost \\
        Insured Losses (\$) & MeanLosses & SDLosses & MinLosses - MaxLosses \\
        \hline
    \end{tabular}
    \caption{Summary Statistics}
    \cite{statista, ncai, fema}
\end{table}

\subsection{Regression Results}
Table 5 \ref{tab:reg_results} presents results from the panel regressions. The number of disasters and total cost are significantly 
associated with higher premiums based on the log-log specification. A 10\% increase in disasters corresponds to a 14.7\% rise in 
premiums. Meanwhile, a 10\% rise in total damage leads to a 23.1\% premium increase.

@@

    \cite{statista, ncai}
  \end{table}

  The results presented in Table 6 \ref{tab:reg_peril} indicate that hurricane and flood disasters have the most 
  substantial impacts on insurance premiums. Specifically, a 10\% increase in hurricane or flood Damages are associated with 
  approximately a 0.01\% increase in premiums. Severe storms and winter weather also exhibit positive effects on premiums. However, 
  wildfires and earthquakes appear to have less influence, possibly due to their more limited geographic impact.

@@

    \cite{statista, ncai}
  \end{table}

>>>>

% double spacing
\usepackage{setspace}
\doublespacing

@@

\section{Introduction}
\label{sec:intro}
The frequency and severity of weather disasters have increased in recent years, posing threats to homeowners' safety and finances. 
This study investigates the relationship between these disasters and property insurance pricing. As climate disasters become more 
severe, homeowners face increased risks, necessitating a closer look at their ability to protect assets through insurance.

The study emphasizes the catastrophic impact of natural disasters on homeowners and the insurance industry. The financial burden 
falls on homeowners, who must manage insurance premiums, deductibles, and coverage limitations. Likewise, insurance companies face 
increased risks, threatening their financial stability.

While previous research has touched on weather risks and insurance prices, our study delves deeper into the dynamics of frequency and 
severity in property insurance pricing, based on valuable insights \cite{hurricaneeco}. We want to find effective ways to reduce the financial burden 
on homeowners during major disasters.

We draw data from reliable sources such as the Insurance Information Agency, National Insurance Centers for Environmental Information, 
and Aon's 2023 Weather, Climate, and Disaster Investigation Report using statistical analysis and data science. This study aims to 
improve homeowner well-being and the stability of the property insurance market in the face of rising weather-related risks.

@@

These data sources, sourced from leading insurance analytics firms and research organizations, provide an important foundation and 
context for assessing the relationship between intensifying extreme weather events and rising property insurance costs. Using decades 
of specialized knowledge to track climate impacts and pricing trends across vulnerable geographies.

Aon is at the forefront of risk analysis in the insurance industry, with dedicated catastrophe model development and annual global 
climate reports tracking disaster losses. This establishes Aon as an unrivaled resource for linking increasing weather perils to 
property insurance pricing trends.

The Insurance Information Institute (III), a trusted insurance trade group, provides proprietary research and public education on aligning 
property insurance costs with escalating climate risks confronting communities across risk-prone regions.

The National Centers for Environmental Information (NCEI) provides data backbone to guide evidence-based assessment of how record-breaking 
weather extremes are translating to property insurance market dynamics by maintaining the nation's authoritative disaster cost 
database.

The National Association of Insurance Commissioners (NAIC) facilitates transparency from insurers on pricing and product shifts caused by 
worsening climate factors such as hurricanes, flooding, and wildfires through regulatory oversight across states, structured data 
calls, and cross-collaborations.

@@

As evidenced by this chart \ref{fig:10economic_loss} of the ten most expensive global insured catastrophe losses from 1900 to 2022, 
the rising frequency and severity of extreme weather events in the last two decades has increased the risks and costs for property 
insurers. Hurricanes that have ravaged North America and the Caribbean in recent years have caused almost all of the disasters, with 
seven of the ten occurring in the last 15 years. The confluence of billion-dollar hurricane disasters, as well as rapidly rising 
nominal and inflation-adjusted insured losses, highlight the insurance industry's mounting climate-change costs. With three of the 
four largest loss events occurring in 2021-2022 alone, property insurers face escalating climate threats and exposures that 
necessitate significant adjustments to rates, coverages, and risk mitigation efforts in especially vulnerable coastal and island 
regions. Addressing these challenges through preventative resilience initiatives and innovative insurance solutions will be critical 
to enabling the industry to cover the property landscape of the future in a viable and sustainable manner.

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.8\linewidth]{AON Top 10 Cyclones Economic Loss.pdf}
    \label{fig:10economic_loss}
    \cite{aon}
\end{figure}

This chart \ref{fig:10global_loss} of the top ten most economically destructive tropical cyclones in the world from 1900 to 2022 emphasizes the growing risks 
that hurricanes pose in a changing climate. With the United States being hit by all ten of the costliest storms since 2000, the trend 
toward repeated devastating hurricane seasons is clear and concerning for coastal communities. The clustering of five massive 
back-to-back hurricane events in the last six years, putting unprecedented strain on the US economy, lends credence to modeling that 
suggests tropical storms are becoming more destructive. Hurricane vulnerability and recovery costs will only increase as sea levels 
rise and waters warm, exposing more property than ever before.  This necessitates immediate action across the public and private 
sectors to strengthen pre-disaster mitigation and post-disaster resilience if the already record-breaking cost of hurricanes is not 
to skyrocket in the future. To address this escalating climate threat, innovative policy, infrastructure, community design, and 
financing tools must be combined.

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.8\linewidth]{AON Top 10 Global Insured Loss.pdf}
    \label{fig:10global_loss}
    \cite{aon}
\end{figure}

@@

In order to isolate the relationships between disaster activity and average homeowners insurance premiums over time, we used 
panel regression techniques with state and year-fixed effects. To reduce omitted variable bias, models specifically control for 
time-invariant differences across states (via state dummies) as well as national trends (via year indicators). The premium measure is 
the dependent variable, with the key independent variables being disaster counts and cost totals.

These disaster measures capture state-specific exposure each year, providing granular, longitudinal pricing connectivity. Taking logs 
accounts for skewness and simplifies elasticity interpretation. We anticipate that worsening loss shocks will have a direct impact on 
insurers' underwriting decisions, resulting in measurable premium impacts. Differential sensitivity testing is possible by segmenting 
by peril type.

@@

\section{Results}
\label{sec:resu}
\subsection{Summary Statistics}
Table 4 \ref{tab:summary} presents summary statistics for the premium and disaster data. The disaster measures show substantial variability, 
highlighting the irregular nature of extremes. Hurricane and flooding perils account for the largest share of overall cost and 
insured losses. The regression results show that increasing natural disaster activity plays a statistically significant role in 
driving increases in homeowners insurance premiums across risk-exposed states. According to the elasticity estimates, premiums grow 
at a higher percentage rate than disaster costs.

\begin{table}[h]
    \label{tab:summary}
    \centering
    \begin{tabular}{|l|c|c|c|}
        \hline
        & Mean & SD & Range \\
        \hline
        Premium & \$\num{959.2} & \$\num{238.5208} & \$\num{536} - \$\num{1311} \\
        Number of Disasters & MeanDisaster & SDDisaster & MinDisaster - MaxDisaster \\
        Total Cost (\$) & MeanCost & SDCost & MinCost - MaxCost \\
        Insured Losses (\$) & MeanLosses & SDLosses & MinLosses - MaxLosses \\
        \hline
    \end{tabular}
    \caption{Summary Statistics}
    \cite{statista, ncei, FEMA}
\end{table}

\subsection{Regression Results}
Table 5 \ref{tab:reg_results} presents results from the panel regressions. The number of disasters and total cost are significantly 
associated with higher premiums based on the log-log specification. A 10\% increase in disasters corresponds to a 14.7\% rise in 
premiums. Meanwhile, a 10\% rise in total damage leads to a 23.1\% premium increase. In other words, 10\% increase in total catastrophe 
damages incurred in a state-year, for example, corresponds to a 23.1\% increase in annual average premiums. This emphasizes the 
exponentially rising underwriting risks and loss adjustments made by insurers based on compounding extreme event data.

@@

    \cite{statista, ncei}
  \end{table}

  The results presented in Table 6 \ref{tab:reg_peril} indicate that hurricane and flood disasters have the most 
  substantial impacts on insurance premiums. Specifically, a 10\% increase in hurricane or flood Damages are associated with 
  approximately a 0.01\% increase in premiums. Results also show that hurricanes and flooding disasters have a disproportionate 
  impact in comparison to other hazards. Climate change appears to be worsening exposure along the coast and in low-lying areas. 
  Additional findings indicate that winter storms may necessitate price changes in some states. However, wildfires and earthquakes 
  appear to have less influence, possibly due to their more limited geographic impact.

@@

    \cite{statista, ncei}
  \end{table}

  These panel regressions provide empirical evidence that worsening extremes are increasing premiums in line with rising expected loss 
  trends. Strategic risk-reduction policies will be critical in the future.
